---
title: "fasrsfunctions Read Me"
author: "John Walker"
date: `r sys.Date'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##### *Functions for processing US traffic accident data*
The `farsfunctions` package assists in reading and summarizing of US Department of Transportation data compiled by the National
Highway Traffic Safety Administration (NHTSA) provided in the Fatality Analysis Reporting System (FARS). See the references 
section below for access to data, data descriptions and much more about FARS.

## Package installation

The latest version of the `farsfunctions` package is installed from GitHub as shown here: 
```{r install, eval=FALSE}
#Can skip this line if devtools is already installed
install.packages("devtools")

library(devtools)
install_github("jrwalker-projects/farsfunctions") #installs the package
library(farsfunctions) #loads package into R
```

Use of the functions is described in the package vignette.

[![Travis build status](https://travis-ci.org/jrwalker-projects/farsfunctions.svg?branch=master)](https://travis-ci.org/jrwalker-projects/farsfunctions)
  
