---
title: "Using farsfunctions"
author: "John Walker"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using farsfunctions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

##### *Support functions for rocessing US traffic accident data*
The `farsfunctions` package assists in reading and summarizing of US Department of Transportation data compiled by the National
Highway Traffic Safety Administration (NHTSA) provided in the Fatality Analysis Reporting System (FARS). See the references 
section below for access to data, data descriptions and much more about FARS.

## Installation

The latest version of the `farsfunctions` package is installed from GitHub as shown here: 
```{r install, eval=FALSE}
#Can skip this line if devtools is already installed
install.packages("devtools")

library(devtools)
install_github("jrwalker-projects/farsfunctions") #installs the package
library(farsfunctions) #loads package into R
```

## Accident data files

```{r setwd}
#setwd(system.file("extdata",package="farsfunctions"))
#knitr::opts_knit$set(root.dir = normalizePath(".."))
knitr::opts_knit$set(root.dir = system.file("extdata",package="farsfunctions"))
dir()
getwd()
```

```{r dir}
dir()
getwd()
```

## Using the farsfunctions 

### Reading accident files for a set of years

### Summarizing accidents

### Plot a map of accident locations for a US state in a given year

## References
For file downloads and more information on FARS data see
[the main FARS website:](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars)

as of October 2017 see [the latest manual on FARS data:](https://crashstats.nhtsa.dot.gov/Api/Public/Publication/812499)

or see the current document list [NCSA publications](https://crashstats.nhtsa.dot.gov/#/DocumentTypeList/22)

## Acknowledgements
None of the R functions were written by me. They were provided by person(s) unknown as part of the Coursera learning
platform as part of the Johns-Hopkins course "R Packages". My only role as part of the coursework was to document the functions,
create an R package and show that the package can be built and deployed. All of the code, documentation and packaging is 
available on GitHub at the location shown in the Installation section above.
